Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-15 main.tex
:pencil: Update propositions

% amsthm
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}

@@

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{U}^{n \times n}$ denotes the set of $n \times n$ unitary matrices.
	$\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	$\mathrm{tr}(\cdot)$ and $\det(\cdot)$ evaluates the trace and determinant of a square matrix, respectively.
	$\mathrm{diag}(\cdot)$ constructs a square matrix with arguments on the main diagonal and zeros elsewhere.
	$\mathrm{sv}(\cdot)$ returns the singular value vector.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ is the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$ $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, $r$-th iterated point, and final solution, respectively.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\lvert \cdot \rvert$ denotes the absolute value.
	% $\lVert \cdot \rVert _p$ means the $p$-norm and $\lVert \cdot \rVert$ suggests $p = 2$.
	$\lVert \cdot \rVert$ means the Euclidean norm.
	$\lVert \cdot \rVert _\mathrm{F}$ represents the Frobenius norm.
	$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}

@@

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,\min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.

@@

		% The Pareto singular value problem \eqref{op:pareto} can thus be solved by Algorithm~\ref{ag:rcg}.
		\begin{lemma}\label{lm:pareto_gradient}
			The Euclidean gradient of \eqref{ob:pareto} w.r.t. \gls{bd}-\gls{ris} group $g$ is
			\begin{equation}
				\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
				\label{eq:pareto_gradient}
			\end{equation}
			where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		\end{lemma}
		\begin{proof}
			Let $\mathbf{H} = \sum_n \mathbf{u}_n \sigma_n \mathbf{v}_n^\mathsf{H}$ be the compact \gls{svd} of the equivalent channel.
			Since the singular vectors are orthonormal, the $n$-th singular value can be expressed as
			\begin{equation}
				\sigma_n = \mathbf{u}_n^\mathsf{H} \mathbf{H} \mathbf{v}_n = \mathbf{u}_n^\mathsf{T} \mathbf{H}^* \mathbf{v}_n^*,
			\end{equation}
			whose differential w.r.t. $\mathbf{\Theta}_g^*$ is
			\begin{align*}
				\partial \sigma_n
				& = \partial \mathbf{u}_n^\mathsf{T} \underbrace{\mathbf{H}^* \mathbf{v}_n^*}_{\sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T} \mathbf{v}_n} + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \underbrace{\mathbf{u}_n^\mathsf{T} \mathbf{H}^*}_{\mathbf{u}_n^\mathsf{T} \sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T}} \partial \mathbf{v}_n^*\\
				& = \underbrace{\partial \mathbf{u}_n^\mathsf{T} \mathbf{u}_n^*}_{\partial 1 = 0} \cdot \sigma_n + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \sigma_n \cdot \underbrace{\mathbf{v}_n^\mathsf{T} \partial \mathbf{v}_n^*}_{\partial 1 = 0}\\
				& = \mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^* \cdot \mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\\
				& = \mathrm{tr}(\mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^*).
			\end{align*}
			According to \cite{Hjorungnes2007}, the corresponding complex derivative is
			\begin{equation}
				\frac{\partial \sigma_n}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{u}_n \mathbf{v}_n^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
				\label{eq:derivative_sv}
			\end{equation}
			Linear combination of \eqref{eq:derivative_sv} yields \eqref{eq:pareto_gradient}.
		\end{proof}

@@

	\begin{subsection}{Some Analytical Bounds}\label{ss:bounds}
		We then discuss some analytical bounds related to channel singular values.
		\begin{proposition}[rank-deficient channel]\label{pp:dof}
			In point-to-point \gls{mimo}, \gls{bd}-\gls{ris} cannot achieve a higher \gls{dof} than diagonal \gls{ris}.
		\end{proposition}
		\begin{proof}
			The scattering matrix of \gls{bd}-\gls{ris} can be decomposed as\footnote{This is because (block) unitary matrices are closed under multiplication.}
			\begin{equation}
				\mathbf{\Theta} = \mathbf{L} \mathbf{\Theta}_\mathrm{D} \mathbf{R}^\mathsf{H},
			\end{equation}
			where $\mathbf{\Theta}_\mathrm{D} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$ corresponds to diagonal \gls{ris} and $\mathbf{L}, \mathbf{R} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$ are block diagonal matrices of $L \times L$ unitary blocks.
			Manipulating $\mathbf{L}$ and $\mathbf{R}$ rotates the linear spans of $\bar{\mathbf{H}}_\mathrm{B} \triangleq \mathbf{H}_\mathrm{B} \mathbf{L}$ and $\bar{\mathbf{H}}_\mathrm{F} \triangleq \mathbf{R}^\mathsf{H} \mathbf{H}_\mathrm{F}$ and maintains their rank.
			On the other hand, there exists a $\mathbf{\Theta}_\mathrm{D}$ such that
			\begin{equation*}
				\begin{split}
					\mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{\Theta}_\mathrm{D} \mathbf{H}_\mathrm{F})
					& = \min \bigl( \mathrm{rank}(\mathbf{H}_\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}_\mathrm{D}), \mathrm{rank}(\mathbf{H}_\mathrm{F}) \bigr) \\
					& = \min \bigl( \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{B}), N_\mathrm{S}, \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{F}) \bigr) \\
					& = \max_\mathbf{\Theta} \ \mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F})
				\end{split}
			\end{equation*}
			The same result holds if the direct link is present.
		\end{proof}

		\begin{proposition}[\gls{los} forward\footnote{A similar result holds for \gls{los} backward channel.} channel]\label{pp:los}
			If the forward channel is rank-1, then \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the $n$-th ($n \ge 2$) channel singular value to the $(n-1)$-th (resp. $n$-th) singular value of $\mathbf{T}$, that is,
			\begin{equation}
				\sigma_1(\mathbf{T}) \ge {\sigma_2(\mathbf{H})} \ge \sigma_2(\mathbf{T}) \ge \ldots \ge \sigma_{N-1}(\mathbf{T}) \ge {\sigma_N(\mathbf{H})} \ge \sigma_N(\mathbf{T}),
				\label{iq:sv_bound_los}
			\end{equation}
			where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{v}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ and $\mathbf{v}_\mathrm{F}$ is the right compact singular vector of $\mathbf{H}_\mathrm{F}$.
			Note that $\sigma_1(\mathbf{H})$ is unbounded with a sufficiently large \gls{ris}.
		\end{proposition}
		\begin{proof}
			Let $\mathbf{H}_\mathrm{F} = \sigma_\mathrm{F} \mathbf{u}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}$ be the compact \gls{svd} of the forward channel.
			The channel Gram matrix can be written as Hermitian-plus-rank-1:
			\begin{equation}
				\mathbf{G} \triangleq \mathbf{H} \mathbf{H}^\mathsf{H} = \mathbf{Y} + \mathbf{z} \mathbf{z}^\mathsf{H},
			\end{equation}
			where $\mathbf{Y} \triangleq \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{v}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ and $\mathbf{z} \triangleq \sigma_\mathrm{F} \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{u}_\mathrm{F} + \mathbf{H}_\mathrm{D} \mathbf{v}_\mathrm{F}$.
			By the Cauchy interlacing formula \cite{Golub2013}, the $n$-th ($n \ge 2$) eigenvalues of $\mathbf{G}$ are bounded by
			\begin{equation}
				\lambda_1(\mathbf{Y}) \ge \lambda_2(\mathbf{G}) \ge \lambda_2(\mathbf{Y}) \ge \ldots \ge \lambda_{N-1}(\mathbf{Y}) \ge \lambda_N(\mathbf{G}) \ge \lambda_N(\mathbf{Y}).
				\label{iq:ev_bound_los}
			\end{equation}
			Since $\mathbf{Y} = \mathbf{T} \mathbf{T}^\mathsf{H}$ is positive semi-definite, the eigenvalues are non-negative and taking the square root of \eqref{iq:ev_bound_los} gives \eqref{iq:sv_bound_los}.
		\end{proof}
		It is worth notice that a similar conclusion holds for diagonal \gls{ris} \cite{Semmler2023}.
		We will later show that for a finite $N_\mathrm{S}$, using a larger group size can approach those bounds better.

		\begin{proposition}[fully-connected \gls{ris} without direct link]
			% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
			If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then
			\begin{equation}
				\mathrm{sv}(\mathbf{H}) = \mathrm{sv}(\mathbf{BF}),
			\end{equation}
			where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
		\end{proposition}

		\begin{proof}
			Let $\mathbf{H}_\mathrm{B} = \mathbf{U}_\mathrm{B} \mathbf{\Sigma}_\mathrm{B} \mathbf{V}_\mathrm{B}^\mathsf{H}$ and $\mathbf{H}_\mathrm{F} = \mathbf{U}_\mathrm{F} \mathbf{\Sigma}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
			The scattering matrix of fully-connected \gls{ris} can be decomposed as
			\begin{equation}
				\mathbf{\Theta} = \mathbf{V}_\mathrm{B} \mathbf{X} \mathbf{U}_\mathrm{F}^\mathsf{H},
				\label{eq:scattering_fc}
			\end{equation}
			where $\mathbf{X} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$ is a unitary matrix to be designed.
			The equivalent channel is thus a function of $\mathbf{X}$
			\begin{equation}
				\mathbf{H} = \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{U}_\mathrm{B} \mathbf{\Sigma}_\mathrm{B} \mathbf{X} \mathbf{\Sigma}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}.
				\label{eq:channel_equivalent_fc}
			\end{equation}
			Since $\mathrm{sv}(\mathbf{U} \mathbf{A} \mathbf{V}^\mathsf{H}) = \mathrm{sv}(\mathbf{A})$ for unitary $\mathbf{U}$ and $\mathbf{V}$, we have
			\begin{align*}
				\mathrm{sv}(\mathbf{H}) & = \mathrm{sv}(\mathbf{U}_\mathrm{B} \mathbf{\Sigma}_\mathrm{B} \mathbf{X} \mathbf{\Sigma}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H})\\
				& = \mathrm{sv}(\mathbf{\Sigma}_\mathrm{B} \mathbf{X} \mathbf{\Sigma}_\mathrm{F})\\
				& = \mathrm{sv}(\bar{\mathbf{U}}_\mathrm{B} \mathbf{\Sigma}_\mathrm{B} \mathbf{\bar{V}}_\mathrm{B}^\mathsf{H} \bar{\mathbf{U}}_\mathrm{F} \mathbf{\Sigma}_\mathrm{F} \mathbf{\bar{V}}_\mathrm{F}^\mathsf{H})\\
				& = \mathrm{sv}(\mathbf{BF}),
			\end{align*}
			where $\bar{\mathbf{U}}_{\mathrm{B}/\mathrm{F}}$ and $\bar{\mathbf{V}}_{\mathrm{B}/\mathrm{F}}$ are arbitrary unitary matrices.
		\end{proof}

>>>>

% amsthm
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[proposition]

@@

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{H}^{n \times n}$ and $\mathbb{U}^{n \times n}$ denotes the set of $n \times n$ Hermitian and unitary matrices, respectively.
	% $\mathbb{U}^{n \times n}$ denotes the set of $n \times n$ unitary matrices.
	$\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	$\mathrm{tr}(\cdot)$ and $\det(\cdot)$ evaluates the trace and determinant of a square matrix, respectively.
	$\mathrm{diag}(\cdot)$ constructs a square matrix with arguments on the main diagonal and zeros elsewhere.
	$\mathrm{sv}(\cdot)$ returns the singular value vector.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ is the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$ $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, $r$-th iterated point, and final solution, respectively.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\lvert \cdot \rvert$ denotes the absolute value.
	% $\lVert \cdot \rVert _p$ means the $p$-norm and $\lVert \cdot \rVert$ suggests $p = 2$.
	$\lVert \cdot \rVert$ means the Euclidean norm.
	$\lVert \cdot \rVert _\mathrm{F}$ represents the Frobenius norm.
	$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}

@@

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.

@@

		% The Pareto singular value problem \eqref{op:pareto} can thus be solved by Algorithm~\ref{ag:rcg}.
		\begin{lemma}\label{lm:pareto_gradient}
			The Euclidean gradient of \eqref{ob:pareto} w.r.t. \gls{bd}-\gls{ris} group $g$ is
			\begin{equation}
				\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
				\label{eq:pareto_gradient}
			\end{equation}
			where $\mathbf{U}$ and $\mathbf{V}$ are the left and right singular matrices of $\mathbf{H}$, respectively.
		\end{lemma}
		\begin{proof}
			Please refer to Appendix~\ref{ap:pareto_gradient}.
		\end{proof}

@@

	\begin{subsection}{Some Analytical Bounds}\label{sc:bounds}
		We then discuss some analytical bounds related to channel singular values.
		\begin{proposition}[degree of freedom]\label{pp:dof}
			In point-to-point \gls{mimo}, \gls{bd}-\gls{ris} cannot achieve a higher \gls{dof} than diagonal \gls{ris}.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:dof}.
		\end{proof}

		\begin{proposition}[rank-deficient channel]\label{pp:rank_deficient}
			% If the forward/backward channel is rank-$k$, then a \gls{bd}-\gls{ris} can at most enlarge the $n$-th ($n > k$) channel singular value to the $n-k$-th singular value of $\mathbf{T}$, or suppress the $n$-th ($n < N - k + 1$) channel singular value to the $n$-th singular value of $\mathbf{T}$.
			% If the forward/backward channel is rank-$k$, then a sufficiently large \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ equivalent channel singular values without bounds.
			If the forward or backward channel is rank-$k$, then regardless of the \gls{ris} size and architecture, the $n$-th singular value of the equivalent channel is bounded by
			\begin{align}
				\sigma_n(\mathbf{H}) & \le \sigma_{n-k}(\mathbf{T}), && \text{if } n > k, \label{iq:sv_bound_enlarge}\\
				\sigma_n(\mathbf{H}) & \ge \sigma_n(\mathbf{T}), && \text{if } n < N - k + 1, \label{iq:sv_bound_suppress}
			\end{align}
			where
			% \begin{align}
			% 	\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
			% 	\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k.
			% \end{align}
			\begin{equation}
				\mathbf{T} \mathbf{T}^\mathsf{H} =
				\begin{cases}
					\mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
					\mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{B} \mathbf{U}_\mathrm{B}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k,
				\end{cases}
			\end{equation}
			and $\mathbf{V}_\mathrm{F}$ and $\mathbf{U}_\mathrm{B}$ are the right and left compact singular matrices of $\mathbf{H}_\mathrm{F}$ and $\mathbf{H}_\mathrm{B}$, respectively.
			% where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the forward channel is rank-$k$ and $\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the backward channel is rank-$k$.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:rank_deficient}.
		\end{proof}

		% \begin{corollary}[degree of freedom]
		% 	In point-to-point \gls{mimo} with or without direct link, \gls{bd}-\gls{ris} cannot provide a higher \gls{dof} than diagonal \gls{ris}.
		% \end{corollary}

		\begin{corollary}[extreme singular values]
			With a sufficiently large \gls{ris}, the first $k$ channel singular values are unbounded from above while the last $k$ channel singular values can be suppressed to zero.
			% For a rank-$k$ forward or backward channel with a sufficiently large \gls{ris}, the first $k$ singular values of the equivalent channel are unbounded from above while the last $k$ singular values can be suppressed to zero.
			% If the forward or backward channel is rank-$k$ and the \gls{ris} is sufficiently large, then the first $k$ singular values of the equivalent channel are unbounded from above while the last $k$ singular values can be suppressed to zero.
			% If the forward/backward channel is rank-$k$, then the first (resp. last) $k$ singular values of the equivalent channel are unbounded from above (resp. below).
			% a sufficiently large \gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ singular values of the equivalent channel without bounds.
		\end{corollary}

		\begin{corollary}[\gls{los} channel]
			If the forward or backward channel is \gls{los}, then a \gls{ris} can at most enlarge (resp. suppress) the $n$-th ($n \ge 2$) channel singular value to the $(n-1)$-th (resp. $n$-th) singular value of $\mathbf{T}$, that is,
			\begin{equation}
				\sigma_1(\mathbf{T}) \ge {\sigma_2(\mathbf{H})} \ge \sigma_2(\mathbf{T}) \ge \ldots \ge \sigma_{N-1}(\mathbf{T}) \ge {\sigma_N(\mathbf{H})} \ge \sigma_N(\mathbf{T}).
				\label{iq:sv_bound_los}
			\end{equation}
		\end{corollary}

		% \begin{proposition}[\gls{los} forward\footnote{A similar result holds for \gls{los} backward channel.} channel]\label{pp:los}
		% 	If the forward channel is rank-1, then \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the $n$-th ($n \ge 2$) channel singular value to the $(n-1)$-th (resp. $n$-th) singular value of $\mathbf{T}$, that is,
		% 	\begin{equation}
		% 		\sigma_1(\mathbf{T}) \ge {\sigma_2(\mathbf{H})} \ge \sigma_2(\mathbf{T}) \ge \ldots \ge \sigma_{N-1}(\mathbf{T}) \ge {\sigma_N(\mathbf{H})} \ge \sigma_N(\mathbf{T}),
		% 		\label{iq:sv_bound_los}
		% 	\end{equation}
		% 	where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{v}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ and $\mathbf{v}_\mathrm{F}$ is the right compact singular vector of $\mathbf{H}_\mathrm{F}$.
		% 	Note that $\sigma_1(\mathbf{H})$ is unbounded with a sufficiently large \gls{ris}.
		% \end{proposition}
		% \begin{proof}
		% 	Let $\mathbf{H}_\mathrm{F} = \sigma_\mathrm{F} \mathbf{u}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}$ be the compact \gls{svd} of the forward channel.
		% 	The channel Gram matrix can be written as Hermitian-plus-rank-1:
		% 	\begin{equation}
		% 		\mathbf{G} \triangleq \mathbf{H} \mathbf{H}^\mathsf{H} = \mathbf{Y} + \mathbf{z} \mathbf{z}^\mathsf{H},
		% 	\end{equation}
		% 	where $\mathbf{Y} \triangleq \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{v}_\mathrm{F} \mathbf{v}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ and $\mathbf{z} \triangleq \sigma_\mathrm{F} \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{u}_\mathrm{F} + \mathbf{H}_\mathrm{D} \mathbf{v}_\mathrm{F}$.
		% 	By the Cauchy interlacing formula \cite{Golub2013}, the $n$-th ($n \ge 2$) eigenvalues of $\mathbf{G}$ are bounded by
		% 	\begin{equation}
		% 		\lambda_1(\mathbf{Y}) \ge \lambda_2(\mathbf{G}) \ge \lambda_2(\mathbf{Y}) \ge \ldots \ge \lambda_{N-1}(\mathbf{Y}) \ge \lambda_N(\mathbf{G}) \ge \lambda_N(\mathbf{Y}).
		% 		\label{iq:ev_bound_los}
		% 	\end{equation}
		% 	Since $\mathbf{Y} = \mathbf{T} \mathbf{T}^\mathsf{H}$ is positive semi-definite, the eigenvalues are non-negative and taking the square root of \eqref{iq:ev_bound_los} gives \eqref{iq:sv_bound_los}.
		% \end{proof}
		% It is worth notice that a similar conclusion holds for diagonal \gls{ris} \cite{Semmler2023}.
		% We will later show that for a finite $N_\mathrm{S}$, using a larger group size can approach those bounds better.
		% We will later
		In Section~\ref{sc:simulation}, we will show by simulation that a finite-size \gls{bd}-\gls{ris} can approach those bounds better than diagonal \gls{ris}.

		% \begin{proposition}[fully-connected \gls{ris} without direct link]\label{pp:fully_connected}
		% 	% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
		% 	If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then
		% 	\begin{equation}
		% 		\mathrm{sv}(\mathbf{H}) = \mathrm{sv}(\mathbf{BF}),
		% 	\end{equation}
		% 	where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
		% \end{proposition}

		\begin{proposition}[fully-connected \gls{ris} without direct link]\label{pp:fully_connected}
			% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
			If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the channel singular values can be manipulated up to
			% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the bounds on the channel singular values are equivalent to the singular values of the equivalent channel, that is,
			% only bounds that apply here are those that apply to the singular values of
			\begin{equation}
				\mathrm{sv}(\mathbf{H}) = \mathrm{sv}(\mathbf{BF}),
			\end{equation}
			where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
			The problem now becomes how the singular values of matrix product are bounded by the singular values of its individual factors.
			% Our focus thus shifts to how the singular values of matrix product are bounded by the singular values of its individual factors.
		\end{proposition}

		\begin{proof}
			Please refer to Appendix~\ref{ap:fully_connected}.
		\end{proof}

		% We now seek to relate the singular values of the matrix product to the singular values of its factors.

		% The problem now becomes, what can we say about the singular values of the matrix product in terms of the singular values of the factors?

		\begin{corollary}[generic singular value bounds \cite{Fulton2000}]
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{H}) \le \prod_{i \in {I}} \sigma_i(\mathbf{H}_\mathrm{B}) \prod_{j \in {J}} \sigma_j(\mathbf{H}_\mathrm{F}),
				\label{iq:sv_bound_fc}
			\end{equation}
			for all admissible triples $(I, J, K) \in T_r^n$ with $r < n$, where
			\begin{equation}
				\begin{gathered}
					% T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \text{for all } p < r \text{ and all } (F, G, H) \text{ in } T_p^r,\\
					% \sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
					T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \forall p < r, (F, G, H) \in T_p^r,\\
					\sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
				\end{gathered}
			\end{equation}
			\begin{equation}
				U_r^n \triangleq \Bigl\{(I, J, K) \mid \sum_{i \in I} i + \sum_{j \in J} j = \sum_{k \in K} k + r(r+1)/2\Bigr\}.
			\end{equation}
		\end{corollary}
